Observing Earth as an Exoplanet with \TESS: Photometric Mapping of Surface Features in Reflected Light

### 2019-05-03 tex/ms.tex
minor edits

However, obtaining a map of a temperate terrestrial exoplanet is still
beyond the capabilities of current facilities, as both the thermal and 
reflected light signals from these planets
are orders of magnitude below current measurement capabilities. Nevertheless,
much work has been done on the theoretical front on mapping terrestrial
planets in the habitable zone
\citep[e.g.,][]{Ford2001,Tinetti2006a,Tinetti2006b,WilliamsGaidos2008,Palle2008,
Kawahara2010,Fujii2012,Berdyugina2017,Haggard2018,Farr2018,Lustig-Yaeger2018};
for a review, see \cite{Cowan2018}. In particular, several works
have harnessed space-based observations of the Earth in proof-of-concept
studies aimed at shedding light on the opportunities and challenges future
exoplanet mapping missions may bring. Perhaps the most well-known of these
is the work of \cite{Sagan1993}, who used observations from the Galileo probe
to infer the existence of life on Earth. However, this study made use of both
spectroscopic observations and spatially-resolved images of the planet at high
resolution. \cite{Goode2001} and \cite{Woolf2002} used unresolved measurements of
Earthshine reflecting off the face of the Moon to infer the planet's
seasonally-variable albedo and its atmospheric transmission spectrum,
respectively. Similarly, \cite{Hasinoff2011} presented a method to infer
the two-dimensional map of the Earth from single-pixel measurements of the
planet's diffuse reflectance on the Moon.
\cite{Cowan2009} used unresolved images of the Earth taken
by the EPOXI spacecraft to construct low resolution albedo maps, recovering
approximate longitudinal distributions of oceans continents on the planet. 
More recently, \cite{Jiang2018} made use of daily images of the fully-illuminated
disk of the Earth taken by the Deep Space Climate Observatory (DSCOVR) to recover
the planet's rotation rate and seasonal albedo variations.

@@

Although the Earthshine signal in \TESS is convolved with complicated systematics,
this dataset offers an unprecedented opportunity to tackle the mapping problem, given
its extremely high cadence (observations recorded every 2 minutes in short cadence mode) 
and the long observing baseline of \TESS (a 2 year primary mission, with the Earth in view
for several months of the year).
%
Moreover, recent advances on the modeling side 
have made the reconstruction of detailed maps more feasible than ever. 
\cite{Luger2019} developed an analytic algorithm for generating---and
inverting---light curves of stars and planets whose surface emission is
described in terms of spherical harmonics, dubbed \starry. The analyticity of this
method makes the computation of the model both fast and precise.
In turn, \cite{Haggard2018} derived analytic expressions for the
analogous case in reflected light, packaged into the 
\textsf{EARL} code. In this work, we derive similar expressions within
the \starry framework to perform fast, analytical inversion of the light curve
of the Earth and thus construct a map of the planet in reflected light.

@@

\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/map_L2.pdf}
    \caption{\label{fig:map_L2}
             Maximum likelihood map of the Earth obtained during the first
             stage of our optimization procedure, in which an iterative
             analytic approach was used to solve the bilinear model, 
             subject to very conservative regularization. Despite the low
             resolution of this map, bright features appear at the 
             expected locations of South America, Africa, and Asia, 
             and two dark features appear at locations consistent with
             the Pacific and Atlantic oceans. Note that the colors correspond
             to a \emph{normalized} albedo, scaled such that the dynamic
             range of the image spans [0, 1]. \codelink{earthshine_S1_S2}      
             }
    \end{centering}
\end{figure}

@@

\begin{figure}[p!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/map.pdf}
    \caption{\label{fig:map}
             \emph{Top:} The maximum likelihood model for the surface albedo
             of the Earth, weighted by the visibility across Sectors 1 and 2. The
             dashed white lines indicate latitudes above/below which the
             visibility drops below 50\%. As in Figure~\ref{fig:map_L2}, coastlines 
             are overplotted as black contours.
             \emph{Bottom:} Approximate cloud coverage map based on corrected
             surface reflectance obtained by the
             \emph{VIIRS} imager, averaged over the same date range and weighted
             by the same visibility function.
             \codelink{earthshine_S1_S2}
             }
    \end{centering}
\end{figure}

@@

First, many of the issues arise because
of the finite signal-to-noise of our measurements and various biases in
our adopted model. While the measurement errors on our data points are
quite small---since the signal of the Earth is both quite strong and independently
measured at 95 different positions on the detector---our lack of understanding
of the complex optics (and our
attempt to na{\"i}vely model it with pixel-level decorrelation) likely
introduces significant correlated noise when we de-trend. We discuss this
and other sources of bias in more detail in \S\ref{sec:bias} below.

@@

Third, inspection of Figure~\ref{fig:phases} shows that most of the time, the Earth
is close to full phase as seen by \TESS.
With the terminator close to the limb
of the planet, its ability to break degeneracies is considerably lessened.
We are therefore somewhat limited in our ability to infer in particular the
latitudinal position of features. Observations of the Earth in upcoming
sectors may contain longer segments in which the planet is seen in crescent,
which should help mitigate these degeneracies.

@@

It is difficult to avoid being disingenuous in one's solution
when the answer (or the expected answer)
to a problem is known ahead of time. Since we had access to 
``ground truth'' from the beginning, many of the de-trending/processing/fitting
choices we made were (however inadvertently) driven by the desire to
come as close as possible to the correct answer. While many of these choices
correspond to numbers we have explicitly quoted here, such as the prior
variances on our model terms, several of them are implicit decisions on
\emph{how} we performed the analysis. For instance, in an early version of 
the project we used as regressors actual \TESS housekeeping variables---such as the angle 
and elevation of the Earth and the Moon relative to each of the four camera 
boresights---instead of the PLD basis. However, the systematics model constructed from these quantities
proved to be too inflexible, judging in part from the poor fidelity of the recovered
maps. The results in this work
must therefore be taken with a grain of salt. That said, we have attempted to
offset the possibility of confirmation bias by making our model
(Equation~\ref{eq:model}) as simple as possible in the end (a multiplicative baseline
and a linear reflectance model) and thus limiting
the number of knobs we could turn. We also opted to forego any use of 
real imaging data in our model; for instance, we fit for the spectral
slope of the map as a free parameter rather than using the true power spectrum 
of the Earth. Moreover, the broad features in the strongly
regularized map (Figure~\ref{fig:map_L2})---and in particular the monsoon clouds
over southeast Asia---are insensitive to most of our model choices.

>>>>

Obtaining a map of a temperate terrestrial exoplanet is still
beyond the capabilities of current facilities, as both the thermal and 
reflected light signals from these planets
are orders of magnitude below current measurement capabilities. Nevertheless,
much work has been done on the theoretical front on mapping terrestrial
planets in the habitable zone
\citep[e.g.,][]{Ford2001,Tinetti2006a,Tinetti2006b,WilliamsGaidos2008,Palle2008,
Kawahara2010,Fujii2012,Berdyugina2017,Haggard2018,Farr2018,Lustig-Yaeger2018};
for a review, see \cite{Cowan2018}. In particular, several works
have harnessed space-based observations of the Earth in proof-of-concept
studies aimed at shedding light on the opportunities and challenges future
exoplanet mapping missions may bring. Perhaps the most well-known of these
is the work of \cite{Sagan1993}, who used observations from the Galileo probe
to infer the existence of life on Earth. This study made use of both
spectroscopic observations and spatially-resolved images of the planet at high
resolution, neither of which will necessarily be available for an analogous exoplanet. 
For a more exoplanet-like approach, \cite{Goode2001} and \cite{Woolf2002} used unresolved measurements of
Earthshine reflecting off the face of the Moon to infer the planet's
seasonally-variable albedo and its atmospheric transmission spectrum,
respectively. Similarly, \cite{Hasinoff2011} presented a method to infer
the two-dimensional map of the Earth from single-pixel measurements of the
planet's diffuse reflectance on the Moon.
\cite{Cowan2009} used unresolved images of the Earth taken
by the EPOXI spacecraft to construct low resolution albedo maps, recovering
approximate longitudinal distributions of oceans continents on the planet. 
More recently, \cite{Jiang2018} made use of daily images of the fully-illuminated
disk of the Earth taken by the Deep Space Climate Observatory (DSCOVR) to recover
the planet's rotation rate and seasonal albedo variations.

@@

Although the Earthshine signal in \TESS is convolved with complicated systematics,
this dataset offers an unprecedented opportunity to tackle the mapping problem in detail, given
its extremely high cadence (observations recorded every 2 minutes in short cadence mode) 
and the long observing baseline of \TESS (a 2 year primary mission, with the Earth in view
for several months of the year).
%
Moreover, recent advances on the modeling side 
have made the reconstruction of detailed maps more feasible than ever. 
\cite{Luger2019} developed an analytic algorithm for generating---and
inverting---light curves of stars and planets whose surface emission is
described in terms of spherical harmonics, dubbed \starry. The analyticity of this
method makes the computation of the model both fast and precise.
In turn, \cite{Haggard2018} derived analytic expressions for the
analogous case in reflected light, packaged into the 
\textsf{EARL} code. In this work, we derive similar expressions within
the \starry framework to perform fast, analytical inversion of the light curve
of the Earth and thus construct a map of the planet in reflected light.

@@

\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/map_L2.pdf}
    \caption{\label{fig:map_L2}
             Maximum likelihood map of the Earth obtained during the first
             stage of our optimization procedure, in which an iterative
             analytic approach was used to solve the bilinear model, 
             subject to very conservative regularization. Despite the low
             resolution of this map, bright features appear at the 
             expected locations of South America, Africa, and Asia, 
             and two dark features appear at locations consistent with
             the Pacific and Atlantic oceans. The colors correspond
             to a \emph{normalized} albedo, scaled such that the dynamic
             range of the image spans [0, 1]. \codelink{earthshine_S1_S2}      
             }
    \end{centering}
\end{figure}

@@

\begin{figure}[p!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/map.pdf}
    \caption{\label{fig:map}
             \emph{Top:} The maximum likelihood model for the surface albedo
             of the Earth, weighted by the visibility across Sectors 1 and 2 and
			 normalized. The
             dashed white lines indicate latitudes above/below which the
             visibility drops below 50\%. As in Figure~\ref{fig:map_L2}, coastlines 
             are overplotted as black contours.
             \emph{Bottom:} Approximate cloud coverage map based on corrected
             surface reflectance obtained by the
             \emph{VIIRS} imager, averaged over the same date range and weighted
             by the same visibility function.
             \codelink{earthshine_S1_S2}
             }
    \end{centering}
\end{figure}

@@

First, many of the issues arise because
of the finite signal-to-noise of our measurements and various biases in
our adopted model. While the measurement errors on our data points are 
small---since the signal of the Earth is both quite strong and independently
measured at 95 different positions on the detector---our lack of understanding
of the complex optics (and our
attempt to na{\"i}vely model it with pixel-level decorrelation) likely
introduces significant correlated noise when we de-trend. We discuss this
and other sources of bias in more detail in \S\ref{sec:bias} below.

@@

Third, inspection of Figure~\ref{fig:phases} shows that most of the time, the Earth
is close to full phase as seen by \TESS.
With the terminator close to the limb
of the planet, its ability to break degeneracies is considerably lessened.
We are therefore somewhat limited in our ability to infer in particular the
latitudinal position of features. The Earthshine present in upcoming
sectors may contain longer segments in which the planet is seen in crescent,
which should help mitigate these degeneracies.

@@

It is difficult to avoid being disingenuous in one's solution
when the answer (or the expected answer)
to a problem is known ahead of time. Since we had access to 
``ground truth'' from the beginning, many of the de-trending/processing/fitting
choices we made were (however inadvertently) driven by the desire to
come as close as possible to the correct answer. While many of these choices
correspond to numbers we have explicitly quoted here, such as the prior
variances on our model terms, several of them are implicit decisions on
\emph{how} we performed the analysis. For instance, in an early version of 
the analysis we used as regressors actual \TESS housekeeping variables---such as the angle 
and elevation of the Earth and the Moon relative to each of the four camera 
boresights---instead of the PLD basis. However, the systematics model constructed from these quantities
proved to be too inflexible, judging in part from the poor fidelity of the recovered
maps. The results in this work
must therefore be taken with a grain of salt. That said, we have attempted to
offset the possibility of confirmation bias by making our model
(Equation~\ref{eq:model}) as simple as possible in the end (a multiplicative baseline
and a linear reflectance model) and thus limiting
the number of knobs we could turn. We also opted to forego any use of 
real imaging data in our model; for instance, we fit for the spectral
slope of the map as a free parameter rather than using the true power spectrum 
of the Earth. Moreover, the broad features in the strongly
regularized map (Figure~\ref{fig:map_L2})---and in particular the monsoon clouds
over southeast Asia---are insensitive to most of our model choices.
